---
title: "Assignment 3"
author: "Zachariah Alex"
date: "2022-10-17"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
#LOADING LIBRARY FUNCTIONS

library(caret)
library(class)
library(e1071)
library(dplyr)

```

```{r}
#READING CSV FILE

data1<-read.csv("UniversalBank.csv",header=TRUE)
head(data1)

#CONVERTING TO FACTORS 

data1$Personal.Loan <- as.factor(data1$Personal.Loan)
data1$Online <- as.factor(data1$Online)
data1$CreditCard <- as.factor(data1$CreditCard)

is.factor(data1$Personal.Loan)
is.factor(data1$Online)
is.factor(data1$CreditCard)

```


```{r}
#PARTITIONING DATA TO 60:40 RATIO

set.seed(350)
data_partition<-createDataPartition(data1$Personal.Loan,p=.6,list=FALSE,times=1)

train<-data1[data_partition,]
valid<-data1[-data_partition,]
head(train)


```

```{r}

#NORMALIZING THE DATA

norm <- preProcess(train[,-c(10,13,14)], method=c("center","scale"))

train_norm <-predict(norm,train)

valid_norm<- predict(norm,valid)

head(valid_norm)
head(train_norm)

```


\textcolor{red}{A: Creating a pivot table for the training data with Online as a column variable,}
\textcolor{red}{CC as a row variable, and Loan as a secondary row variable}

```{r}

table_loan<-table(train_norm$CreditCard,train_norm$Personal.Loan,train_norm$Online)
View(table_loan)
```
\textcolor{red}{B: The probability of loan acceptance conditional on 
having a bank credit card}
\textcolor{red}{and being an active user of online banking services:}

$$(Loan=1,CC = 1,Online=1)=(53 /(468+53))=0.1017274
$$


\textcolor{red}{C:Creating separate pivot tables for the training data with Loan (rows) as a 
function of Online (columns)}

```{r}
table_1<-table(train_norm$Personal.Loan,train_norm$Online)
View(table_1)
```



\textcolor{red}{C: Creating separate pivot tables for the training data with Loan (rows) as a function of CC.}


```{r}

table_2<-table(train_norm$Personal.Loan,train_norm$CreditCard)
View(table_2)
```

\textcolor{red}{Creating Personal Loan table}

```{r}
table_3<-table(train_norm$Personal.Loan)
proptable_3<-prop.table(table_3)
View(proptable_3)
```


\textcolor{red}{D:Computing the following probabilities from the table}



$${i. P(CC = 1 | Loan = 1)  =84/288=0.2916666}$$

$${ii. P(Online = 1 | Loan = 1)= 180/288=0.625}$$



$${iii. P(Loan = 1)  =288/3000=0.096}$$



$${iv. P(CC = 1 | Loan = 0)  =804/(1908+804)=804/2712=0.2964601}$$




$${v. P(Online = 1 | Loan = 0) =1593/(1593+1119)=1593/2712=0.5873893}$$



$${vi. P(Loan = 0) =2712/3000=0.904}$$



\textcolor{red}{E: Using the quantities computed above and computing the naive Bayes probability }

\textcolor{red}{P(Loan = 1 | CC= 1, Online = 1)}

$$P(Loan=1|CC=1,Online=1) = $$
$$P(CC=1|Loan=1 )* P(Online=1|Loan=1).P(Loan=1))$$
$$\text{----------------------------------------------------------------------------------------------------------- }$$
$${P(CC=1|Loan=1) * P(Online=1|Loan=1).P(Loan=1))+P(CC=1|Loan=0) * }$$
$${P(Online=1|Loan=0) *P(Loan=0) ) }$$


$$( 0.2916*0.625*0.096)/((0.2916*0.625*0.096)+(0.2964*0.5873*0.904))= 0.01014$$


$$P(Loan=1|CC=1,Online=1) =0.01014 $$


\textcolor{red}{F: Comparing this value with the one obtained from the pivot table in B}
$$\text{From pivot table (B)}$$

$${(Loan=1,CC = 1,Online=1)=(53 /(468+53))=0.1017}$$


$$\text{From computing the naive Bayes probability}$$

$$P(Loan=1|CC=1,Online=1) =0.01014 $$
$$\text{We can say that the probability from the pivot table is more accurate as we are directly}$$
$$\text{taking values from the frequency table while  naive bayes probability calculation}$$
$$\text{is based upon assumptions.}$$

\textcolor{red}{G: Running Naive bayes theorem-Using the quantities computed above to compute the naive Bayes probability }

$$P(Loan = 1 | CC = 1, Online = 1)$$
```{r}

model<-naiveBayes(Personal.Loan~CreditCard+Online,data=train_norm)

model


``` 

$$\text{Comparing naiveBayes Probability values with the calculations  ,we can say that this is more accurate than step E.}$$
$$\text{In step E there is a possibility of manual calculation errors and roundoff errors and assumptions.}
$$










