---
title: "FML FINAL PROJECT"
author: "Zachariah Alex"
date: "2022-11-21"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(Warning = FALSE)
```

```{r}
#loading library functions

library(dplyr)
library(caret)
library(factoextra)
library(cluster)
library(esquisse)

```

```{r}
#reading data file

data1<-read.csv("fuel.csv")

```

```{r}
#replacing missing/blank data with NA

data2<-data1%>% replace(.=="", NA)

```

```{r}
#finding average null values of missing data in each column

Null_values<-data2%>%is.na()%>%colMeans()*100
```

```{r}
#removing columns having more than 50% null values and filtering required columns

data4<- subset(data2,select=c(rowid,plant_id_eia,fuel_received_units,fuel_mmbtu_per_unit,sulfur_content_pct,ash_content_pct,mercury_content_ppm,fuel_cost_per_mmbtu,contract_type_code,energy_source_code,fuel_type_code_pudl,fuel_group_code,supplier_name,primary_transportation_mode_code,plant_id_eia_label, natural_gas_transport_code,contract_type_code))


```

```{r}

#Sampling 2% of the data

set.seed(8250)

data5<-sample_n(data4,12000)

```

```{r}

#splitting data to 75:25 test train ratio

set.seed(8578)

sample<-createDataPartition(data5$rowid,p=0.75, list=FALSE)

train<-data5[sample,]

test<-data5[-sample,]

```

```{r}
#combining required categorical and numerical variables

qdata<-train[,c(2,3,4,5,6,7,8,11)]

#replacing NA values with 0 for calculations

cdata<-qdata%>% replace(.=="", NA)

cdata[is.na(qdata)] <- 0

```

```{r}
#Assigning dummy variables to categorical variable fuel_type_code_pudl

fuel_coal <- ifelse(cdata$fuel_type_code_pudl=="coal" ,1,0)
fuel_gas <- ifelse(cdata$fuel_type_code_pudl=="gas" ,1,0)
fuel_oil <- ifelse(cdata$fuel_type_code_pudl=="oil" ,1,0)

data6<-cbind(cdata[,-c(8)],fuel_coal,fuel_gas,fuel_oil)


```


```{r}
#normalizing the data

data8<-scale(data6)


```

*Applying hierarchical clustering algorithm*
```{r}
# Creating the dissimilarity matrix for data set through Euclidean distance.

distance <- dist(data8, method = "euclidean")
```

```{r}
# Hierarchical clustering using Ward's method

clusterfuel <- hclust(distance, method = "ward.D2" )
```
*Wards distance is used because Ward's minimum variance criterion minimizes the total within-cluster variance*

```{r}
# Plotting the dendrogram

plot(clusterfuel, cex = 0.6, hang = -1)
rect.hclust(clusterfuel,k=3,border=2:5)
```
*Cut-off height = 135.Therefore number of clusters =3 *

*Here we are choosing k value =3 (domain knowledge) so as to find the distribution of 3 types of fuel in each of the clusters*


```{r}
#cutting the dendrogram tree for k=3

grp <- cutree(clusterfuel, k = 3)

```

```{r}
# Finding number of members in each of the clusters

table(grp)

```

*Based on the number of clusters considered, 3 groups have been formed with 4891,3338,771 data points*

```{r}
#Binding clusters to main data 

data7<- cbind(data6, clustering = grp)

```

```{r}

#Cluster visualization 

fviz_cluster(list(data = data7, cluster = grp))


```

*The three clusters are named as  Coal, Gas and Oil*

*Cluster 1= GAS*

*Cluster 2= COAL*

*Cluster 3= OIL*

```{r}

#Finding mean of required columns for interpretation

comb_data<-cbind(data7,train[,c(9,11,12,14,15,16)])

fuel<-comb_data %>% mutate(clusters=comb_data$clustering) %>% group_by(clusters)

data9<-fuel[,c(2:11)]%>%group_by(clustering)%>%summarise_all("mean")

```

*Plotting clusters vs other variables*
```{r}

#Cluster vs fuel type 

ggplot(fuel, aes(x=fuel$clusters,fill=fuel$fuel_type_code_pudl))+geom_bar()
```

\newpage
```{r}

#cluster vs heat content

ggplot(data9, aes(x=clustering, y=fuel_mmbtu_per_unit,fill=clustering)) + geom_bar(stat="identity") + 
labs(x="clusters", y="fuel heat content")+scale_fill_gradient(low = "#FF0000", high = "#000000") + theme_minimal()
```
\newpage
```{r}

#cluster vs fuel received

ggplot(data9, aes(x=clustering, y=fuel_received_units,fill=clustering)) + geom_bar(stat="identity") + 
labs(x="clusters", y="fuel recieved")+scale_fill_viridis_c(option = "viridis", direction = 1) +
theme_minimal()
```
\newpage
```{r}

#cluster vs fuel cost

ggplot(data9, aes(x=clustering, y=fuel_cost_per_mmbtu,fill=clustering)) + geom_bar(stat="identity") + 
labs(x="clusters", y="fuel_cost")
```
\newpage
```{r}

# cluster vs sulfur content

ggplot(data9, aes(x=clustering, y=sulfur_content_pct,fill=clustering)) + geom_bar(stat="identity") + 
labs(x="clusters", y="sulfur content")+scale_fill_gradient(low = "#FF0000", high = "#000000") + theme_minimal()
```
\newpage
```{r}
#cluster vs ash content

ggplot(data9, aes(x=clustering, y=ash_content_pct,fill=clustering)) + geom_bar(stat="identity") + 
labs(x="clusters", y="ash content")
```
\newpage
```{r}
#cluster vs mercury content

ggplot(data9, aes(x=clustering, y=mercury_content_ppm,fill=clustering)) + geom_bar(stat="identity") + 
labs(x="clusters", y="mercury content")+scale_fill_viridis_c(option = "viridis", direction = 1) +
theme_minimal()
```
\newpage
```{r}
#cluster vs contract type

ggplot(fuel, aes(x=fuel$clusters,fill=fuel$contract_type_code))+geom_bar(position=position_dodge())+facet_wrap(~contract_type_code)
```
\newpage
```{r}
#cluster vs transportation mode

ggplot(fuel, aes(x=fuel$clusters,fill=fuel$primary_transportation_mode_code))+geom_bar(position=position_dodge())+facet_wrap(~primary_transportation_mode_code) 

```
\newpage

***ANALYSIS OF CLUSTERS***



**CLUSTER ONE: GAS**

*Cluster 1 data accounts to little above half or 55.36% of our sample data.*

*Gas is the main fuel type used in this cluster.*

*Average units of gas received is 405035 units which is the highest compared to coal and oil.*

*Gas is free from components like mercury, sulphur and ash.*

*Cost for the gas fuel is 13.04 USD per MMBtu.*

*Pipeline is the transportation method used for this type of fuel.*




**CLUSTER TWO:COAL**

*Cluster two is formed by 3215 observations which totals up to 35.72% of our sampled data.*

*Fuel type used in this cluster is coal.*

*Average units of coal received is 46216 units*

*Average heat content of the fuel is 21.32 which is higher than other two fuels.*

*Coal fuel has an average mercury content of 0.011%,sulfur content of 1.38% and ash content of 9.93%.*

*Coal energy is cheaper when compared with gas and oil and  average fuel cost per MMBtu is 1.70 USD.*




**CLUSTER THREE :OIL**

*Cluster 3 is formed by 802 observations which totals to only 8.9% of our sample data.*

*Oil or Petroleum is the fuel type used*

*Total quantity of fuel received is 6083 units.*

*Fuel heat content is 5.83 units which is greater than gas.*

*This type of fuel is free from mercury and ash content ,but contains  a little of 0.19% of sulphur .*

*Fuel cost per MMBtu is 9.99 USD.*




**Extra Credit**

```{r}

#building the model

model<-lm(data6$fuel_cost_per_mmbtu~.,data=data6)

summary(model)

#Finding variable importance 

varImp(model)

```
*Here we can see that our p value is greater than 5 % and r squared value is too low.*
*This means variables we considered doesnot account to the variability for fuel cost per mmbtu variable*


