---
title: "ASSIGNMENT 4"
author: "Zachariah Alex"
date: "2022-11-04"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r}

#Loading library functions

library(factoextra)
library(dplyr)
library(ggcorrplot)
```

```{r}
#Reading the csv file

data1<-read.csv('Pharmaceuticals.csv')

```

```{r}
#normalizing the data

data2<-data1[,(3:11)]
data3<-scale(data2)

```

```{r}
#running Elbow method for finding optimal value for k

set.seed(356)
fviz_nbclust(data3,kmeans,method="wss")
```

*From the plot it is clear that k =2*

```{r}

#running Silhouette method to find optimal value for k

set.seed(423)
fviz_nbclust(data3,kmeans,method="silhouette")


```

*From the plot it is clear that k =5*

```{r}

#Running kmean on k = 2 from elbow method

k2<-kmeans(data3,centers=2)

k2

#binding data to dataset

data4<-cbind(data3,k2$cluster)

```

```{r}
#Running kmean on k=5 from silhouette method

k5<-kmeans(data3,centers=5)

k5

#binding data to dataset
data5<-cbind(data3,k5$cluster)


```


***(A)INTERPRETATION***

*Here we are using k-mean algorithm with elbow and silhouette method for cluster analysis*

*We can see that by using elbow method optimal number of clusters obtained are 2*
```{r}
#plotting cluster based on elbow method

fviz_cluster(k2,data2,main="2 clusters")
```

*In silhouette method the optimum number of cluster obtained is 5*
```{r}
#plotting cluster based on silhouette method

fviz_cluster(k5,data2,main="5 clusters")
```

***(B)Analysing clusters***

```{r}
#Finding mean of each coumn within each clusters in elbow method

data2 %>% mutate(Cluster = k2$cluster) %>% group_by(Cluster) %>% summarise_all("mean")
```

*From the results we can see that cluster one stocks is good to invest than cluster two*

*Market cap is higher for cluster one and beta value which indicates the volatility is lower for cluster one*

*Optimal range of PE is twenty to twenty five range and cluster one has Price to earning ratio of twenty one*

*Moreover ROE,ROA,Asset turnover and most important net profit margin is higher for cluster one stocks.*

```{r}
#Finding mean of each coloumns in each clusters in silhouette method

data2 %>% mutate(Cluster = k5$cluster) %>% group_by(Cluster) %>% summarise_all("mean")
```

*From the results,we can see that cluster five stocks is good to invest than rest four clusters*

*Market cap is highest and lower beta value(low risk) for cluster 5*

*Optimal range of PE is twenty to twenty five range and cluster one has Price to earning ratio of twenty two*

*Moreover ROE,ROA,Asset turnover and most important net profit margin is highest for cluster five stocks.*


***(C) Pattern in the clusters with respect to the variables (10 to 12)***

```{r}

#BASED ON EXCHANGE 

plot <-data1[12:14] %>% mutate(Clusters=k2$cluster)
ggplot(plot, mapping = aes(factor(Clusters), fill =Exchange))+geom_bar(position='dodge')+labs(x ='Clusters')



plot <-data1[12:14] %>% mutate(Clusters=k5$cluster)
ggplot(plot, mapping = aes(factor(Clusters), fill =Exchange))+geom_bar(position='dodge')+labs(x ='Clusters')


```

*FROM PLOTTING,WE CAN SEE THAT IN BOTH ELBOW METHOD AND SILHOUETTE METHOD IN ALL THE CLUSTERS FORMED MOST OF THE COMPANIES ARE LISTED IN NEWYORK STOCK EXCHANGE*

```{r}
#BASED ON RECOMMENDATION

plot <-data1[12:14] %>% mutate(Clusters=k2$cluster)
ggplot(plot, mapping = aes(factor(Clusters), fill =Median_Recommendation))+geom_bar(position='dodge')+labs(x ='Clusters')



plot <-data1[12:14] %>% mutate(Clusters=k5$cluster)
ggplot(plot, mapping = aes(factor(Clusters), fill =Median_Recommendation))+geom_bar(position='dodge')+labs(x ='Clusters')

```


*FROM PLOTTING WE CAN SEE THAT IN BOTH ELBOW METHOD AND SILHOUETTE METHOD IN ALL THE CLUSTERS FORMED WE CAN SEE THAT THERE IS A LEAST NUMBER OF RECOMMENDATIONS FOR STRONG BUY*
  

```{r}

#BASED ON LOCATION 

plot <-data1[12:14] %>% mutate(Clusters=k2$cluster)
ggplot(plot, mapping = aes(factor(Clusters), fill =Location))+geom_bar(position='dodge')+labs(x ='Clusters')



plot <-data1[12:14] %>% mutate(Clusters=k5$cluster)
ggplot(plot, mapping = aes(factor(Clusters), fill =Location))+geom_bar(position='dodge')+labs(x ='Clusters')


```

*FROM PLOTTING WE CAN SEE THAT IN BOTH ELBOW METHOD AND SILHOUETTE METHOD IN ALL THE CLUSTERS FORMED MOST OF THE COMPANIES ARE BASED IN USA*


***(D) CLUSTER NAMINGS***

*FROM ELBOW METHOD*
 
*CLUSTER 1 -MODEST BULL BASKET (based on all variables-medium buy stocks)*

*CLUSTER 2 -CREEPING BULL BASKET (based on all variables)*

*FROM SILHOUETTE METHOD*

*CLUSTER 1 -BUDDING BULL BASKET (based on revenue growth)*

*CLUSTER 2 -PENNY BEAR BASKET (based on all variables)*

*CLUSTER 3 -HARMONY BULL BASKET (based on all variables-consistent)*

*CLUSTER 4 -SPECULATIVE BULL BASKET (based on PE ratio)*

*CLUSTER 5 -GOLDEN BULL BASKET (based on all variables)*
